\newpage


\addcontentsline{toc}{chapter}{Bibliography}

\begin{thebibliography}{99}
%	\makeatletter
%	\renewcommand{\@biblabel}[1]{}
%	\makeatother

%
%\bibitem{Baldacci}
%R. Baldacci. \textit{Scienza delle Costruzioni}, UTET, 1984.

\bibitem{Abeyaratne} R. Abeyaratne, \textit{Continuum Mechanics},  Lecture Notes, 2012.

\bibitem{Anand}
L. Anand, S. Govindjee. \textit{Continuum Mechanics of solids}, OUP Oxford, 2020.

%\bibitem{Benvenuto}
%	E. Benvenuto {\em La Scienza delle Costruzioni e il suo sviluppo
%		storico}, Manuali Sansoni Firenze, 1981.

%\bibitem{Corradi}
%L. Corradi dell'Acqua, \textit{Meccanica delle strutture vol. 1. Il comportamento dei corpi continui}, McGraw-Hill Education, 2010

\bibitem{Barber}
J.R. Barber. \textit{Elasticity},  Springer, Berlin, Heidelberg, 4th edition, 2022

\bibitem{Carlson}
 D.E. Carlson. \textit{Linear Thermoelasticity}. In: Truesdell, C. (eds) Linear Theories of Elasticity and Thermoelasticity. Springer, Berlin, Heidelberg,  1973.

\bibitem{Gurtin}
M.E. Gurtin. \textit{The Linear Theory of Elasticity}. Handbook of Physics, Vol. VIa/2, Springer, 1972.

\bibitem{Gurtin1}
M.E. Gurtin. \textit{An Introduction to Continuum Mechanics}, Academic Press, 1982.

\bibitem{Gurtin_2003} 
Gurtin, M.E., 2003. On a framework for small-deformation viscoplasticity: free
energy, microforces, strain gradients. International Journal of Plasticity 19, 47--90.

\bibitem{Gurtin2}
M.E. Gurtin, E. Fried, L. Anand. \textit{The Mechanics and Thermodynamics of Continua}, Cambridge University Press, 2010.

%\bibitem{Paolone} A. Luongo, A. Paolone, \textit{Scienza delle costruzioni. Il Problema di de Saint Venant}, CEA, 2005.

\bibitem{Paroni}
R. Paroni. \textit{Scienza delle Costruzioni.  Elementi di teoria dell’elasticità lineare}, Springer, 2022.

\bibitem{Primer}
P. Podio Guidugli. \textit{A Primer in Elasticity}, Kluwer Academic, 2000.

\bibitem{Elageo}
P. Podio Guidugli, A. Favata, \textit{Elasticity for Geotechnicians}, Springer, 2014.

\bibitem{PPG1}
P. Podio Guidugli. \textit{Continuum Thermodynamics}, Springer, 2019.

\bibitem{sadd}
 M.H. Sadd. \textit{Elasticity: Theory, Applications, and Numerics}, Academic Press, 5th edition, 2025

\bibitem{TadmorMiller}
E.B. Tadmor, R.E. Miller. \textit{Modeling Materials: Continuum, Atomistic and Multiscale Techniques}, Cambridge University Press, 2011.

\bibitem{BornHuang}
M. Born, K. Huang. \textit{Dynamical Theory of Crystal Lattices}, Oxford University Press, 1954.

\bibitem{AshcroftMermin}
N.W. Ashcroft, N.D. Mermin. \textit{Solid State Physics}, Holt, Rinehart and Winston, 1976.

\bibitem{KeelerBatchelder1970}
G.J. Keeler, D.N. Batchelder. Measurement of the elastic constants of argon from 3 to 77~K. \textit{Journal of Physics C: Solid State Physics} \textbf{3} (1970), 510.

\bibitem{Reif}
F. Reif. \textit{Fundamentals of Statistical and Thermal Physics}, McGraw-Hill, 1965.

\bibitem{PodioGuidugli2019}
P. Podio-Guidugli. The virial theorem: a pocket primer. \textit{Journal of Elasticity} \textbf{137} (2019), 219--235.

\bibitem{Truesdell1960} C. Truesdell, R. Toupin, \textit{The classical field theories}, in Handbuch der Physik, vol. III/1, 1960.


\bibitem{Truesdell2}
C. Truesdell, W. Noll. \textit{The Nonlinear Field Theories of Mechanics}. In Fl\"{u}gge, S. (Ed.), Handbuch der Physik III/3. Springer, Berlin, 1965.

\bibitem{Truesdell3} C. Truesdell, \textit{The Elements of Continuum Mechanics}, Lectures given in August - September 1965 for the Department of Mechanical and Aerospace Engineering Syracuse University Syracuse, New York.

\bibitem{Truesdell1977} C. Truesdell, \textit{A First Course in Rational Continuum Mechanics}, Academic Press, 1977.

\bibitem{Truesdell1}
C. Truesdell. \textit{Rational Thermodynamics },  Springer, Berlin, Heidelberg,  1984.



%\bibitem{Truesdell}
%C. Truesdell. \textit{A First Course in Rational Continuum Mechanics}, Academic
%Press, 2016.

%\bibitem{Turesdell1968} C. Truesdell, \textit{The nonlinear field theories in mechanics} , pp. 19--215 of Topics in Nonlinear Physics, Berlin-Heidelberg-New York, Springer- Verlag, 1968.



		
%\bibitem{Podio1}
%P. Podio Guidugli {\em Lezioni di Scienza delle Costruzioni}, Aracne editrice, 2009.

%\bibitem{Podio2}
%P. Podio Guidugli {\em Lezioni di Statica}, Aracne editrice, 2014.

%\bibitem{Sollazzo}
%A. Sollazzo, S. Marzano. \textit{Scienza delle Costruzioni} vol. 2, UTET, 1988.

%\bibitem{Vidoli}
%S. Vidoli, M. Gei, \textit{Scienza delle costruzioni. Dalla meccanica dei solidi alle strutture}, Flaccovio. 2024.
		
\end{thebibliography}